### 2024-07-18 code/eda-code/eda.qmd
Further exploration, prepping data for modeling, model tuning and multiple feature set iterations

---
title: "Exploratoray Analysis"
date: today
output: html_document
---

# Setup

```{r}
# load needed packages. make sure they are installed.
pacman::p_load(here, #for data loading/saving
               tidyverse, # for data manipulation and plotting
               skimr, # for data exploration and summary
               ggthemes, # for plot theming
               tidytext) # for text analysis
```

Load the data.

```{r}
# path to data
data_location <- here::here("data","processed-data","processeddata.rds")
# load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here::here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

We are saving the results for each of the plots to the `results` folder.

# Data exploration through figures

Density plots for the continuous outcomes -- email body length first. We'll utilize the `ggplot2` visualization package, which offers a `geom_density` shape that is well suited to visualizing the distributions of multiple groups in the same plot. In our case that will be both spam and legitimate emails.

```{r, warning=FALSE}
p1 <- mydata %>% ggplot(aes(x = bodylength, fill = label)) +
      geom_density(position = "dodge", alpha = 0.5) +
      xlim(0,7500) + xlab("Length") + ylab("Density") +
      labs(title = "Body Distribution", caption = "x-axis limited to a maximum of 7,500", fill = NULL) +
      scale_y_continuous(labels = scales::percent) +
      theme_tufte() +
      theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom")
plot(p1)
figure_file = here::here("results", "figures","bodylength-distribution.png")
ggsave(filename = figure_file, plot = p1) 
```

And we'll produce the same visualization for subject line length.

```{r, warning=FALSE}
p2 <- mydata %>% ggplot(aes(x=subjectlength, fill = label)) +
      geom_density(position = "dodge", alpha = 0.5) +
      xlim(0,150) + xlab("Length") + ylab("Density") +
      labs(title = "Subject Line Distribution", caption = "x-axis limited to a maximum of 150", fill = NULL) +
      scale_y_continuous(labels = scales::percent) +
      theme_tufte() +
      theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom")
plot(p2)
figure_file = here::here("results", "figures","subjectlength-distribution.png")
ggsave(filename = figure_file, plot = p2) 
```
Now the most common sender domains by both spam and legitimate emails. We'll continue to utilize `ggplot2`, this time with the `geom_col` shape in order to produce a bar chart of counts faceted by the label.

```{r}
p3 <- mydata %>% 
        select(label, senderdomain) %>% 
        filter(!senderdomain == "") %>% 
        count(label, senderdomain) %>% 
        group_by(label) %>% 
        slice_max(n, n = 10) %>% 
        ungroup() %>% 
        mutate(senderdomain = reorder(senderdomain, n)) %>% 
        ggplot(aes(n, senderdomain, fill = label)) +
        geom_col() +
        facet_wrap(~label, scales = "free") +
        labs(title = "Most Popular Sender Domains by Type") +
        ylab(NULL) + xlab(NULL) +
        scale_x_continuous(labels = scales::comma) +
        theme_tufte() +
        theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "none")
plot(p3)
figure_file = here::here("results", "figures", "popular-domains.png")
ggsave(filename = figure_file, plot = p3) 
```

```{r}
# plot the most popular email body tokens
p4 <- mydata %>%
        select(label, subject) %>% 
        unnest_tokens(word, subject) %>% 
        anti_join(stop_words) %>% 
        filter(!grepl('[0-9]', word)) %>% 
        count(label, word) %>% 
        group_by(label) %>% 
        slice_max(n, n = 15) %>% 
        ungroup() %>% 
        mutate(word = reorder(word, n)) %>% 
        ggplot(aes(n, word, fill = label)) +
        geom_col() +
        facet_wrap(~label, scales = "free") +
        labs(title = "Most Popular Subject Tokens by Type") +
        ylab(NULL) + xlab(NULL) +
        scale_x_continuous(labels = scales::comma) +
        theme_tufte() +
        theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "none")
plot(p4)
figure_file = here::here("results", "figures", "popular-subject-tokens.png")
ggsave(filename = figure_file, plot = p4)

p5 <- mydata %>%
        select(label, body) %>% 
        unnest_tokens(word, body) %>% 
        anti_join(stop_words) %>% 
        filter(!grepl('[0-9]', word)) %>% 
        count(label, word) %>% 
        group_by(label) %>% 
        slice_max(n, n = 15) %>% 
        ungroup() %>% 
        mutate(word = reorder(word, n)) %>% 
        ggplot(aes(n, word, fill = label)) +
        geom_col() +
        facet_wrap(~label, scales = "free") +
        labs(title = "Most Popular Body Tokens by Type") +
        ylab(NULL) + xlab(NULL) +
        scale_x_continuous(labels = scales::comma) +
        theme_tufte() +
        theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "none")
plot(p5)
figure_file = here::here("results", "figures", "popular-body-tokens.png")
ggsave(filename = figure_file, plot = p5) 
```

Similarly we can plot the most common body tokens and the most common subject tokens to get an idea of the word choice between the two labels.

When plotting subject tokens we can see completely distinct collections of the most popular tokens across labels. This bodes well for subject tokens as a potential feature component to supplement analysis of the email bodies with in our final model.

When plotting body tokens we can see that 'http', which precedes most links within each email, is the most common token for each label. Considering this token should approximate not just the presence of a URL, but also the quantity of URLs within each individual email, it will be interesting to see if including the supplied `URLs` boolean will be of value or not in our end model. We suspect that the overlap will result in unnecessary noise and dimensionality to our final model. Additionally, considering the above bar chart suggests link presence and quantity is very similarly distributed across both labels, it may wind up improving model performance altogether to exclude 'http' or URL features from the final model.

```{r, warning=FALSE}
p6 <- mydata %>% 
        select(label, recursive) %>% 
        count(label, recursive) %>% 
        group_by(label) %>% 
        summarize(percent = n / sum(n)) %>% 
        ungroup() %>% 
        filter(percent < .5) %>% 
        ggplot(aes(label, percent, fill = label)) +
        geom_col() +
        labs(title = "Emails with the Same Sender/Receiver") +
        ylab(NULL) + xlab(NULL) +
        scale_y_continuous(labels = scales::percent) +
        theme_tufte() +
        theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "none")
plot(p6)
figure_file = here::here("results", "figures", "recursive-emails.png")
ggsave(filename = figure_file, plot = p6) 
```

Lastly, we examine an engineered feature called `recursive`. This feature is a boolean for whether or not the sender and receiver emails are the same. A common strategy when sending emails in bulk is to include the sender email in the To line while hiding a large list of recipients within the BCC line. In order to see if this is a strategy employed significantly by spam email senders we plotted the percentage of recursive emails for both legitimate and spam emails. We can see that in both cases the percentages are quite low (<2.5%) with only about a 0.7% difference between the two labels. Based on this, it seems unlikely that this feature will warrant further investigation or inclusion in our final model.